---
title: "Week 2: R Exercise and Homework"
output: github_document
---

# Analysis Routine

On August 29, DJ Patil, the Chief Data Scientist in the White House Office
of Science and Technology Policy, and Mark Rosekind, the Administrator of
the National Highway Traffic Safety Administration (NHTSA) announced the
release of a data set documenting all traffic fatalities occurring in the
United States in 2015. In this excise, we tried to use the command we have
learned to analyze this dataset.

## Analysis Routine: Data Import

- To get started with this exercise, direct your R session to a dedicated
workshop directory and download into this directory the following sample
tables. Then import the files into Excel and save them as tab delimited
text files.

    - [accident](./accident.csv)
    - [Description of the fields in the dataset](ftp://ftp.nhtsa.dot.gov/FARS/FARS-DOC/Analytical User Guide/USERGUIDE-2015.pdf)

- Import the tables into R

```{r import_data1, eval=T}
accidents <- read.delim(file="accident.csv", header=T, sep=",") 
head(accidents)
names(accidents)
``` 
- Preprocess the data

```{r import_data2, eval=T}
#Only include data with correct long/lat coordinates and valid hours
accidents <- subset(accidents, LONGITUD!=999.99990 &  LONGITUD!=888.88880 & LONGITUD!=777.77770 & HOUR<30)
``` 

- Aggregate the number of accidents by month 

```{r import_data3, eval=T}
hist(accidents$MONTH)
``` 

- Aggregate the number of accidents by hour 

```{r import_data4, eval=T}
hist(accidents$HOUR)
``` 

- Aggregate the number of persons by month

```{r import_data5, eval=T}
a<-aggregate(PERSONS~MONTH, accidents,sum)
plot(a$MONTH, a$PERSONS, type='o', col='blue')
``` 

- Aggregate the number of persons by hour

```{r import_data6, eval=T}
a<-aggregate(PERSONS~HOUR, accidents,sum)
plot(a$HOUR,a$PERSONS, type='o', col='blue' )
``` 

# Homework

Based on the `accident` dataset, please write a function (see the
							  [Rbascis.md](./Rbasics.md)),
say named `GetStateFatality(stateName, display)`, to return the monthly
fatalites of at state `stateName` and if `display=TRUE` also to plot the
fatalities by months. The following is the code for each U.S. states and
Puerto Rico:

01 Alabama 31 Nebraska
02 Alaska 32 Nevada
04 Arizona 33 New Hampshire
05 Arkansas 34 New Jersey
06 California 35 New Mexico
08 Colorado 36 New York
09 Connecticut 37 North Carolina
10 Delaware 38 North Dakota
11 District of Columbia 39 Ohio
12 Florida 40 Oklahoma
13 Georgia 41 Oregon
15 Hawaii 42 Pennsylvania
16 Idaho 43 Puerto Rico
17 Illinois 44 Rhode Island
18 Indiana 45 South Carolina
19 Iowa 46 South Dakota
20 Kansas 47 Tennessee
21 Kentucky 48 Texas
22 Louisiana 49 Utah
23 Maine 50 Vermont
24 Maryland 52 Virgin Islands 
25 Massachusetts 51 Virginia
26 Michigan 53 Washington
27 Minnesota 54 West Virginia
28 Mississippi 55 Wisconsin
29 Missouri 56 Wyoming
30 Montana

```{r homework, eval=T}
## Now finish the following R function, starting from directly reading the csv
## file (accidents.csv). 
#install.packages('ggplot2')
library(ggplot2)
pal <- c("#A6CEE3","#2B7FB8","#97CC91","#55B145","#B29C71","#ED5454","#EE6441","#FDA746","#F08C3A","#B89CCB","#9068be","#f8df31")
month <- c('Jan','Feb','Mar','Apr','May','Jun','Jul','Aug','Sep','Oct','Nov','Dec')

GetStateFatality <- function (stateName, display=FALSE){
  if(is.character(stateName)==FALSE){
    paste('    >>>State name should be a string.')
    return()
  }
  if(is.logical(display)==FALSE){
    paste('    >>>The second variable should be longical, that is TRUE or FALSE.')
    return()
  }
  
  state.str <- '01 Alabama 31 Nebraska
02 Alaska 32 Nevada
04 Arizona 33 New Hampshire
05 Arkansas 34 New Jersey
06 California 35 New Mexico
08 Colorado 36 New York
09 Connecticut 37 North Carolina
10 Delaware 38 North Dakota
11 District of Columbia 39 Ohio
12 Florida 40 Oklahoma
13 Georgia 41 Oregon
15 Hawaii 42 Pennsylvania
16 Idaho 43 Puerto Rico
17 Illinois 44 Rhode Island
18 Indiana 45 South Carolina
19 Iowa 46 South Dakota
20 Kansas 47 Tennessee
21 Kentucky 48 Texas
22 Louisiana 49 Utah
23 Maine 50 Vermont
24 Maryland 52 Virgin Islands 
25 Massachusetts 51 Virginia
26 Michigan 53 Washington
27 Minnesota 54 West Virginia
28 Mississippi 55 Wisconsin
29 Missouri 56 Wyoming
30 Montana'
  
  #read state string into a pool maxtix 
  str.sub <- gsub('[\n]',' ',state.str)
  str.num <- unlist(strsplit(str.sub,'[[:alpha:][:space:]]+'))
  str.name <- unlist(strsplit(str.sub,'[[:digit:][:space:]]{3,4}'))[-1]
  unlist(strsplit(str.sub,'[[:digit:][:space:]\t]{3,4}'))
  #str.num
  #length(str.name)
  str.pool <- data.frame(Number=as.numeric(str.num), Name=str.name, stringsAsFactors = FALSE)

  #str.split1 <- unlist(strsplit(state.str,'\n'))

  #match state by name, get state number
  state.ind <- 9999
  state.ind <- which(tolower(str.pool$Name) == tolower(stateName))
  if(length(state.ind)==0){
    paste('    >>>No matching state, please check the input state name.')
    return()
  }
  state.num <- str.pool$Number[state.ind]

  #extract data of matched state
  state <- accidents[which(accidents$STATE == state.num),]
  state.monthorder <- state[order(state$MONTH),]
  monthly.fatal <- aggregate(state.monthorder$FATALS,by = list(state.monthorder$MONTH), FUN = sum)
  names(monthly.fatal) <- c('Month','Fatality')
  monthly.fatal$Month <- factor(month, levels = month, ordered = T)
  print(monthly.fatal)

  if(display == TRUE){
    ggplot(data = monthly.fatal,aes(x=1:length(Month),y=Fatality,fill=Month))+
    geom_bar(stat="identity",position=position_dodge())+
    ggtitle(paste('Month Fatality of ',str.pool$Name[state.ind]))+
    ylab('Fatality')+
    scale_fill_manual(name='Month',values = pal)+
    scale_x_continuous(breaks=1:length(monthly.fatal$Month),labels=month)+
    theme(plot.title = element_text(hjust = 0.5),axis.title.x = element_blank())
  }
}

#### Test your function
GetStateFatality ('Texas', display=TRUE)

``` 




