---
title: "Week 9 Homework"
output: github_document
---

# Question 1
The 'trees2.csv' file in the Data folder is a subset of Joshua tree location we
used for last week's homework and raster measurements of climate conditions
('ext5.grd') for the same area.  Please add your lines codes in the following
snippets of the codes to address the question I put in the comments:


```{r hw8, eval=T, echo=T}
# plot the trees 
library(sp)
library(maptools)
library(maps)
library(mapdata)
library(dismo)
library(rgeos)

locs = read.csv(file="Data/trees2.csv", header=T)

# Generate psuedo-absence tree locations as background data

coordinates(locs)=c('lngitude', 'latitude')
proj4string(locs) = CRS('+init=epsg:4326')

x = circles(coordinates(locs), d=50000, lonlat=T)
bg = spsample(x@polygons, 1000, type='random', iter=1000)


# load the climate conditions raster datasets
clim = brick('Data/ext5.grd')

plot(clim, 1, cex=0.5, legend=T, mar=par("mar"), xaxt="n", yaxt="n", main="Annual mean temperature") 
map("state", xlim=c(-119, -113), ylim=c(33.5, 38), fill=F, col="cornsilk", add=T)

# presence of trees
points(locs, col='red')
# psuedo-absense
points(bg, col='blue')

# extracting bioclim values for locations
presence_bc = extract(clim, coordinates(locs)) # for the presence points
bg_bc = extract(clim, bg) # for the pseudo-absence points

# put all the data including presence locations and related climate conditions into a data frame
presence_bc = data.frame(lon=coordinates(locs)[,1], lat=coordinates(locs)[,2], presence_bc)

# do the same thing for background pseudo-absence data

bg_bc = data.frame(lon=coordinates(bg)[,1], lat=coordinates(bg)[,2], bg_bc)

length(which(is.na(bg_bc$bio1))) # double-check for missing data
bg_bc = bg_bc[!is.na(bg_bc$bio1), ] # and pull out the missing values


group_p = kfold(presence_bc, 5) # divide the presence data frame into 5 groups 
group_a = kfold(bg_bc, 5) # same for bg_bc



## To do this, you need to set up `maxent' method in the `dismo` package.
## The maxent package can be downloaded from: http://biodiversityinformatics.amnh.org/open_source/maxent/. Unzip the downloaded tarball and copy the maxent.jar file to the
# folder specified by the following 'jar'. You will also need to install the 'rJava' package. Then you should be ready to go. Please let me know if you have trouble installing this. 

jar <- paste(system.file(package="dismo"), "/maxent/maxent.jar", sep= '') 
jar

## Till this point, all the data prepreration has finished. Now your turn.
##  Based on the kfold results, choose one group (say group 2) as your
##  validation dataset, and use the other four groups as your training
##  data. Use MaxEnt to generate a presense probability map of Joushua
##  trees, plot it, output the accuracy measure. Please identify the
##  climate conditions that affect the most to the Joushua tree
##  distribution and explain how. The sample codes can be found at the MaxEnt section of https://github.com/surfcao/geog5330-fall2017/blob/master/week9/sdm2.Rmd

library(rJava)

p.train <- presence_bc[group_p != 5, ]
p.valid <- presence_bc[group_p == 5, ]

a.train <- bg_bc[group_a != 5, ]
a.valid <- bg_bc[group_a == 5, ]

xm <- maxent(clim, p = p.train[,1:2], a = a.train[,1:2])

plot(xm)

response(xm)
```

According to the contribution plot above, bio8 has the most significant effect on the distribution of the Joshua tree.


```{r}
ev <- evaluate(p.valid[,1:2], a.valid[,1:2], xm, clim)
ev

px <- predict(clim, xm, progress='')
par(mfrow=c(1,2))
plot(px, main = 'Maxent, raw values')

th <- threshold(ev,'spec_sens')
plot(px > th, main = 'Presence/Absence')
points(p.train[,1:2], pch = '+')

```

# Question 2

Please sign up on Google Earth Engine at
[https://signup.earthengine.google.com/#/](https://signup.earthengine.google.com/#/). We
will need it for later part of the class, and it may take a while for
Google to approve your request.




