--
title: "Homework: Use R as GIS"
output: github_document
---
Based on the accident dataset ('accident.dbf') we used last week, please
write R scripts to finish the following two tasks:

1. Map the locations of accidents based on the lat/long and use the United
   States state boundary as the background (the boundary shapefile is in
   the folder 'Data/US'). The result map should have the basic cartographic
   elements (e.g., map scale, compass). Please note the `accident.dbf` can
   be loaded by the following codes, and there are invalid values for
   lat/long column that you probably need to clean up before use (see the
   example code in the homework \#2).  

```{r load, eva=T, echo=T, results='hide'}
#***rm(list=ls()) remove all environment variables
x <- c("sp", "rgdal", "rgeos", "maptools", "classInt", "RColorBrewer", "GISTools", "maps", "raster", 'ggmap')
#install.packages(x) # warning: this may take a number of minutes
lapply(x, library, character.only = TRUE) #load the required packages

library(foreign)
accident=read.dbf('Data/accident.dbf')

# Step 1: Clean up the data first
accident <- subset(accident, LONGITUD!=999.99990 &  LONGITUD!=888.88880 & LONGITUD!=777.77770 & HOUR<30)


# Step 2: Map 
fatality <- accident$FATALS

nclr<- 8
fatalclr<-brewer.pal(nclr, "Reds")
class<-classIntervals(fatality, nclr, style="fixed")
clocode<-findColours(class, fatalclr)

usmap <- readOGR("./Data/US","US_Bndry_Jursidictional_TIGER") #read polygon shapefile
#usmap <- readShapePoly("Data/US/US_Bndry_Jursidictional_TIGER.shp")

plot(usmap,xlim=c(-180,-60),axes=TRUE, col=alpha("gray70", 0.3)) #plot US shapefile

#add title, scalebar, north arrow, and legend
#coordinates(accident) <- c("LONGITUDE","LATITUDE")
#points(accident, pch=19, col=clocode, cex=0.08)
points(accident$LONGITUD, accident$LATITUDE, pch=19, col=clocode, cex=0.08) #add fatality locations on top of Lubbock block shapefile
title(main="Fatality Distribution of United States") 

legend("topright", title = "Occurance", legend=unique(fatality), fill =attr(clocode, "palette"), cex=0.5, bty="n")
north.arrow(xb=-85, yb=65, len=1, lab="N")
```

2. In the folder 'Data/Texas' contains the boundary of Texas counties.
   Using the `accident.dbf' data, please count the number of accidents for
   each county and make a county map based on the value of count you
   calculated. 

```{r texas, eva=T, echo=T}
#install.packages("climdiv")
library(plyr)
library(foreign)
library(ggplot2)
accident=read.dbf('Data/accident.dbf')
state <- accident[which(accident$STATE == 48),]
county.fatal <- aggregate(state$FATALS,by = list(state$COUNTY), FUN = sum)
names(county.fatal) <- c('ID','Count')

#generate color palette
n <- 8
county.col <- brewer.pal(n, "Reds")
county.class <- classIntervals(county.fatal$Count,n,style = "quantile")
colcode<-findColours(county.class, county.col)

txmap <- readOGR("./Data/texas","Tx_CntyBndry_Jurisdictional_STRATMAP")
county.id <- txmap@data$FIPS

county.col <- rep('#FFFFFF',length(county.id))

for (i in 1:length(county.fatal$ID)){
  ind <- which(county.id == county.fatal$ID[i])
  county.col[ind] <- colcode[i]
}

# Use the function  *over()*; see the example codes for the case of bears and parks
# in the lecture

plot(txmap,axes=TRUE, col=county.col)
title(main="Distribution of Accidents in Texas") 
legend("topright", title = "Occurance sum", legend=c("0",names(attr(colcode, "table"))), fill =c("#FFFFFF",attr(colcode, "palette")), cex=0.5, bty="n")
north.arrow(xb=-110, yb=34, len=0.5, lab="N")


# txmap@data$col <- as.factor(county.col)
# txmap@data$id <- rownames(txmap@data)
# txmap.fort <- fortify(txmap, region = 'id')
# txmap.df <- join(txmap.fort,txmap@data,by='id')
# 
# ggplot(data = txmap.df, aes(x = long, y = lat, fill=id, group = group)) +
#   geom_polygon() +
#   geom_path(color="white") +
#   coord_equal() 
#     geom_polygon(colour = "black") +
#     coord_equal() +
#     theme()
```

