---
title: "Week 8 Homework"
output: github_document
---

The `trees.txt` file in the included `Data` folder includes presence locations
of Joshua trees in the southern California, Nevada and Arizona. Please add your
lines codes in the following snippets of the codes to address the question I put
in the comments:


```{r hw7, eval=T, echo=T}
# plot the trees 
library(sp)
library(maptools)
library(maps)
library(mapdata)
library(spatstat)
library(rgdal)
locs = read.csv(file="Data/trees.txt", header=T, sep="\t")

# First to look at the data 
data(stateMapEnv)
plot(c(-119, -112), c(33.0, 38), mar=par("mar"), xlab="longitude", ylab="latitude", xaxt="n", yaxt="n", type="n", main="Joshua tree presence data")

map("state", xlim=c(-119, -112), ylim=c(33.0, 38), fill=T, col="cornsilk", add=T)
points(locs$lngitude,locs$latitude)


# change map projection of original lat/lon data.

coordinates(locs)=c('lngitude', 'latitude')
proj4string(locs) = CRS('+init=epsg:4326')

projLocs<-spTransform(locs, CRS("+init=epsg:2163"))

# Here, please add codes to generate the kernel density estimation map
# of the trees.
locs.ppp <- ppp(projLocs@coords[,1],projLocs@coords[,2],c(min(projLocs@coords[,1]),max(projLocs@coords[,1])),c(min(projLocs@coords[,2]),max(projLocs@coords[,2])))
locs.den <- density(locs.ppp)
plot(locs.den, main='density')
plot(locs.ppp, add = TRUE, cex = 0.5, pch = 20)

# Please repeat the above KDE codes using the following options of
# `sigma` parameters and describe the difference: `sigma=10`,
# `sigma=1000` and `sigma=10000`
locs.den1 <- density(locs.ppp,sigma = 10)
locs.den2 <- density(locs.ppp,sigma = 1000)
locs.den3 <- density(locs.ppp,sigma = 10000)
plot(locs.den1, main='density with sigma of 10')
plot(locs.den2, main='density with sigma of 1000')
plot(locs.den3, main='density with sigma of 10000')
```

In the plots for sigma of 10 and 1000, there are some spikes over the region, while in plot with sigma of 1000, these spikes get smoothened, resulting in a vague density map. 

```{r}
# Please add codes to generate the sample curves of K functions and
# run test to verify whether the data are clustered or not.

library(Hmisc)
Ecdf(pairdist(locs.ppp), main='CDF curve')
plot(Kest(locs.ppp, correction='none'))

```

With the plot result shown above, the data is clustered.