---
title: "Week 4: Homework"
output: github_document
---

```{r global_options, echo=FALSE, results='asis'}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',  warning=FALSE, message=FALSE)
rm(list=ls())
library(sp) 
library(gstat) 
library(RColorBrewer) 
library(plotGoogleMaps)
```
# Question 1

On a 100x100 raster grid, please generate a set of (10 per set)
unconditional simulation maps for each of the following sets of variogram
parameters:

1. sill= 0.5, range = 5, nugget = 0.0001, model = Exponential 
2. sill= 0.5, range = 5, nugget = 0.0001, model = Gaussian 
3. sill= 0.5, range = 50, nugget = 0.0001, model = Exponential 

Based on what you generated, please describe the major visual differences
of the simulated results and explain why. 

```{r q1, echo=T, eval=T}
library(gstat)

# See the section of uncontional simulation in the example codes in the lecture (https://github.com/surfcao/geog5330-fall2017/blob/master/week6/kriging.Rmd)

# create a empty grid
xy <- expand.grid(1:100, 1:100)
names(xy) <- c('x','y')

# Now comment out the following line of code and make the changes of the variogram parameters according to the questions. 

testVgm1 <- vgm(psill=0.5, range=5, model='Exp', nugget=0.0001)
testVgm2 <- vgm(psill=0.5, range=5, model='Gau', nugget=0.0001)
testVgm3 <- vgm(psill=0.5, range=50, model='Exp', nugget=0.0001)

# The following codes are used to run stochastic simulation based on the specified variogram
g.dummy <- gstat(formula=z~1, locations=~x+y, dummy=T, beta=1, model=testVgm1, nmax=21) 
yy <- predict(g.dummy, newdata=xy, nsim=10)
coordinates(yy)=~x+y
gridded(yy) =TRUE
spplot(yy)

g.dummy <- gstat(formula=z~1, locations=~x+y, dummy=T, beta=1, model=testVgm2, nmax=21) 
yy <- predict(g.dummy, newdata=xy, nsim=10)
coordinates(yy)=~x+y
gridded(yy) =TRUE
spplot(yy)

g.dummy <- gstat(formula=z~1, locations=~x+y, dummy=T, beta=1, model=testVgm3, nmax=21) 
yy <- predict(g.dummy, newdata=xy, nsim=10)
coordinates(yy)=~x+y
gridded(yy) =TRUE
spplot(yy)
# Now, please comment on the differences of the different variogram models:

```

When comparing dummy maps from model 1 and 2, they have similar object sizes, but model 2 performs a relatively smmoth pattern while model 1 has a sharp pattern. This difference may come from different models generating the variogram curve. 

For comparison between map 1 and map 3, they have different object sizes, specifically the dummy map 3 has larger size, which leads to larger homogeneity through the region. This phenomenon is from different range values of the variogram model, which corresponds to the size of the surface object.

# Question 2

The `oa.shp` in the `Data` folder is a set of wells containing information
about water level (`WTE`) in feet. Using the `WTE' data for the year of
2013 (field `WTE2013`), please finish the following questions:

1. please interpolate the wells into a continuous surface of water table
   level using simple kriging method and display the estiamted surface.
   Note that `0` in `WTE2013` means no values for those wells.  
   
```{r q21, eval=T, echo=T}
library('sp')
library('maptools')
library('gstat')
library('raster')
wells<-readShapePoints('Data/oa.shp')
wells<-subset(wells,wells$WTE2013>0)
# explore the data by plotting them out
plot(wells)
hist(wells$WTE2013, breaks=20)

# finish the codes for sample variogram and fit the variogram
well.vgm <- variogram(WTE2013 ~ 1,wells)
plot(well.vgm)

well.fit <- fit.variogram(well.vgm, model = vgm(4000, "Gau", 25000, 10))
plot(well.vgm,well.fit)

# The following lines create an empty grid that can be used in krige() function

wellsExtent=bbox(wells)
xy <- expand.grid(seq(wellsExtent[1,1], wellsExtent[1,2],143), seq(wellsExtent[2,1], wellsExtent[2,2],143))
names(xy) <- c('x','y')
coordinates(xy)=~x+y
gridded(xy)=T


# Now finish the codes for kriging and plot the prediction out
well.kriged = krige(wells$WTE2013~1, wells, xy, model = well.fit)
plot(well.kriged[,1], main='Prediction of water level in 2013, simple kriging estimation')

# Plot the interpolated surface using kriging, overlayed by the wells
points(wells,col='red',pch=16)

```
 

2. please also output the map of variance associated with the estimated
   surface and explain what the variance maps tell us. 
   
```{r q22, eval=T, echo=T}

# The output of krige() has two columns, one for prediction, the other for variance. Find the variance and plot it #  out 
plot(well.kriged[,2], main='Variance of water level in 2013, simple kriging estimation')

```

The variance maps can tell us the reliability or accuracy about the predicted or interpolated value at a specific position. In particular, larger variance means less valid interpolation result. And positions with higher variance values should be a reminder that the model could be improved or the source data could have some problems in this place.

3. please run a leave-one-out cross-validation, and calcualte the RMSE
   (root-mean-square deviation) of the estimation. 

```{r q22, eval=T, echo=T}

# For the sake of simplicity, you can use krige.cv() function to get the RMSE; see the section of cross-validation # in the codes of the lecture (https://github.com/surfcao/geog5330-fall2017/blob/master/week6/kriging.Rmd)

cv<-krige.cv(WTE2013 ~ 1, wells, model=well.fit, nfold=nrow(wells)) 
rmse <- sqrt(sum(cv$residual^2))
print(rmse)

```